---
author: "Jimmy Erkens"
title: "AN 588 Homework 1"
date: "`r Sys.Date()`"
output: 
    html_document:
     toc: true
     number_sections: true
     toc_float: true
     code_folding: show
     theme: journal
---

# Libraries 

We first must load and attach R packages that we want to use for the Challenges in Module 4.

```{r, libraries, warning=F}
library(yaml)
library(tidyverse) # install.packages("tidyverse"), literally the best thing ever
```

# Challenges

## Challenge 1

For challenge 1 we take a quote from James Joyce's Ulysses, represented by a vector of length 1, and split it into a vector of several strings. We first remove punctuation using gsub(), then use `str_split()` from `stringr` to remove spaces. This returns a list so we then use `unlist()` to return back to a character vector. Using `seq()` we retain every third element.

```{r, challenge 1}
# I named my vector after Portrait of the Artist of a Young Man because I like it more than Ulysses
potaaym <- c("Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed.")
potaaym <- gsub("[[:punct:]]", "", potaaym) %>% str_split(, pattern = " ") %>% unlist() # %>% is pipe operator, just passes on the data into the next argument
# str_split returns a list, we must unlist in order to select every third item
potaaym <- potaaym[seq(from = 1, to = length(potaaym), by = 3)]; potaaym
```


## Challenge 2

For challenge 2 we do some simple matrix subsetting, fairly self-explainable.

```{r, challenge 2}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE);m

# subset 2nd, 3rd,  6th columns
x <- m[, c(2, 3, 6)]; x

# rows
x <- m[6:8, ]; x

# both rows and columns
x <- m[2:6, 2:9]; x
```


## Challenge 3

For challenge 3 we do some, slightly more complex, array subsetting. Here there are 16 5x5 matrices, indexed by (1:4, 1:4).

```{r, challenge 3}
a <- array(data = order(1:400, decreasing = T), dim = c(5, 5, 4, 4))

a[1, 1, 1, 2]

# this returns the element in the st row, 1st column, of the (1,2) matrix

a[2, 3, 2, ]

# this returns the element in the 3rd row, 2nd column for (1, 2), (2, 2), (3, 2), and (4, 2) matrices

a[1:5, 1:5, 3, 3]

# this returns the entire matrix for the (3, 3) index
```


## Challenge 4

For challenge 4, we create numerous nested lists comprised of character vectors to model taxonomy of _Primates_.

```{r, challenge 4}
# making vectors of each Hap group
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae") #plats
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae") #cats
Tarsioidea <- c("Tarsiidae") #these weirdos

Anthropoidea <- list(Platyrrhini, Catarrhini)
  names(Anthropoidea) <- list("Platyrrhini", "Catarrhini") # nest plats and cats within anthropoids

Haplorhini <- list(Anthropoidea, Tarsioidea)
  names(Haplorhini) <- list("Anthropoidea", "Tarsioidea") # nest anthropoids and tarsiers within haps

Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae", # copy and pasted from online
    "Daubentoniidae") 
Lorisoidea <- c("Lorisidae", "Galagidae") 
Strepsirhini <- list(Lemuroidea, Lorisoidea)
  names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")

Primates <- list(Haplorhini, Strepsirhini)
  names(Primates) <- list("Haplorhini", "Strepsrihini"); Primates # the gang's all here!
```

## Challenge 5

For challenge 5 we filter out a data frame and change the data type for each column. Column 2 gets switched to logical variables, and column 3 gets switched to factor variables.

```{r, challenge 5}
m3 <- matrix(data = c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), nrow = 5, ncol = 3, byrow = T); m3 #beautiful

df_m3 <- data.frame(m3); df_m3
df_m3$X2 <- as.logical(df_m3$X2)
df_m3$X3 <- as.factor(df_m3$X3)
str(df_m3) # we check our work!
```
